EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-07-17 Reply_RefereeA.md

> *On page 3, a large portion of paragraph 3, which introduces DMFT, lacks citations. For example, the authors mention the success of DMFT in capturing key features of various materials, including Mott insulators, heavy fermion compounds, and unconventional superconductors. It would be helpful to include relevant references here to guide interested readers toward further reading.
*

We agree with this comment by the referee. Although it is impossible to do justice to the wide literature accumulated over the years, we have made concerted efforts to incorporate key references that reflect the breadth and depth of research in this field. In the revised version, we have included several citations pertinent to DMFT across the discussed subjects, ensuring that foundational and more recent contributions are acknowledged. This approach aims to provide a balanced overview while addressing the referee's insightful comment.

> *Perhaps a sentence like “Section 6 provides the conclusions” was intended at the end of Section 1 but was inadvertently omitted?*

We sincerely appreciate the referee's prompt attention to this matter. We indeed inadvertently overlooked a reference to Section 6 at the end of the introduction when discussing the manuscript's structure. We resolved this issue in the revised text by including the sentence:   
*"In Sec. 6 we present some concluding remarks and considerations."*

@@

> *Were the spin indices omitted in Eq. (3)? Or do the indices i,j,k,l represent “flavors” that is, combined spin and orbital indices—rather than orbitals alone? The same question applies to the display equation of $H^{int}$ on page 13. This could be confusing to readers, especially since the spin index is also not explicitly shown in the format of the <UMATRIX\_FILE>.restart file described on the same page.*

Thanks for this insightful comment. Indeed, both in Eq.(3) and in the display equation of $H^{int}$ on page 13, the indices $i, j, k, l$  represent "flavors",  corresponding to the combined spin and orbital indices rather than orbitals alone. We acknowledge that this could potentially be confusing, especially given the absence of explicit spin indices in the format of the <UMATRIX\_FILE>.restart file described on the same page.

To clarify this, we have revised the manuscript to explicitly state that the indices denote combined spin and orbital components. Additionally, we will add a note to the relevant sections to highlight this point, ensuring that the readers can interpret the equations and file formats correctly.

We appreciate their attention to this detail and believe that their suggestion will enhance the clarity of our presentation.

@@

> *The last display equation on page 10 seems incorrect, since the creation and annihilation operators do not appear to result in any particle being created or annihilated.*

We express our gratitude to the referee for bringing to our attention the discrepancy in the equations for the bosonic operators. In the revised manuscript, we have rectified this error by modifying the expressions on the right-hand side of the equations to account for the change in the bosonic number.

> *In Sec. 3.3, it appears that users must manually select one of the three provided symmetry configurations to perform block-diagonalization. Is there a plan to extend the code to automatically detect all relevant symmetries—especially in cases where orbitals or sites possess additional symmetries that would allow further block-diagonalization of the matrix?*

We acknowledge the referee’s observation. While the main code currently necessitates manual selection of one of the provided symmetry configurations, we recognize the value of automating the detection of all relevant symmetries. In fact, the Triqs interface ${\tt edipack2triqs}$ features an efficient algorithm to automatically recognize the symmetries of the problem and select the appropriate value for ${\tt ed\_mode}$. 
In the actual version of the Fortran API this is not implemented yet, but we have plans to incorporate it in future release. In the current setup, this requires to accurately analyse the model Hamiltonian and the effective bath decomposition, as discussed in section 3.6, at initialization time.

@@

At the same time, we are actively exploring development avenues that aim to integrate more sophisticated conservation laws, which would further reduce the matrix dimensions. We are confident that these enhancements would improve code’s flexibility and efficiency in managing systems characterized by intricate symmetries.

> *On page 43, it is mentioned that the spectrum obtained from ED is “inherently spiky”. What is the value of the broadening parameter used in these simulations? Including this information in the main text may improve the clarity and reproducibility of the paper.*

We appreciate the referee's  observation. The size of the Lorentzian broadening of the spectral poles is controlled by the input parameter `eps`. It is important to note that the value of `eps` is not fixed universally but rather is chosen based on the specific requirements of each simulation. In the particular examples discussed on page 43 of our manuscript, we have employed a broadening parameter of `eps = 0.01`, expressed in units of the half-bandwidth `D`.
In response to the referee's suggestion, we have updated the manuscript  explicitly specifying the value of this broadening parameter.

### 2025-07-17 Reply_RefereeB.md

> *I find the discussion at the end of page 11 somewhat unclear, especially the terminology. Is this a discussion of consecutive indexing vs. occupation number representation?*

We thank the referee for pointing out this issue. Indeed, the original text aimed to explain how electronic Fock states within a given symmetry sector are indexed and related to their underlying occupation number (bitstring) representation. However, we acknowledge that the terminology and phrasing in the original submission may have been ambiguous.

To clarify this point, we have revised the final paragraph of page 11 in the current version of the manuscript. The updated text now reads:

*“From a computational perspective, constructing a symmetry sector amounts to defining an injective map $\mathcal{M} : \mathcal{S}_{\vec{Q}} \rightarrow \mathcal{F}_e$ that relates the consecutive indexing of the electronic states |i\rangle within a given symmetry sector $\mathcal{S}_{\vec{Q}}$ to the Fock states $|I_e\rangle$ in the electronic Fock space in occupation number representation. This map is typically implemented as a rank-1 integer array, whose size corresponds to the dimension of the sector $D^{\mathcal{S}}$.”*

This new formulation replaces the previous wording and makes explicit the distinction between the consecutive indexing of the sector states $|i\rangle$ and the Fock states in the occupation number representation $|I_e\rangle$. In doing so, we clarify that the mapping $\mathcal{M}$ is what connects the two encodings of the same physical state. We believe the revised text should eliminate the ambiguity the referee noted.

@@

> *The motivation of including the code on p. 12 is unclear. What is the intention here?*

We thank the referee for raising this point. The purpose of the code snippet in question is to illustrate the concrete implementation of the mapping from Fock states to symmetry sectors for different values of ${\tt ed\_mode}$. Our goal was to make explicit the internal logic used to construct the quantum number-conserving subspaces, bridging the gap between formal description and practical realization. We suggest this code snippet is useful to understand the structure of the ${\tt sparse\_map}$ used to implement fast algorithm for the evaluation of the reduced impurity density matrix in Section 3.5.2.

@@

> *p. 16: GFmatrix is said to be a critical component for high-speed execution. Maybe it could be describe in more detail? In what way is it efficient? What does it mean it is multi-layed? In passing, it would be nice if the capitalization would be uniform, e.g. GFmatrix vs. gfmatrix (I understand that Fortran compiler does not care, but the human reader perhaps does).*

We thank the referee for pointing out the need for a clearer and more detailed description of the ${\tt gfmatrix}$ type, which indeed plays a central role in the performance of the code. In response, we have revised the corresponding section of the manuscript (page 16) to provide a more precise explanation of its structure and efficiency features.

The term multi-layered refers to the hierarchical organization of the ${\tt gfmatrix}$ object, which is essential for storing and accessing the analytic structure of correlation functions, i.e. the weights and the poles, in a flexible and scalable way. 
Specifically, the data are organized across three conceptual layers: (1) the many-body eigenstates of the Hamiltonian spectrum, (2) the amplitude channels associated to each such state, and (3) the set of poles and weights corresponding to physical excitations from this state. This structure is necessary to correctly represent the Lehmann decomposition of Green's functions and related objects, while still allowing dynamic access and manipulation during calculations.

The efficiency of ${\tt gfmatrix}$ stems from two main features. First, correlation functions do not need to be precomputed and stored on fixed frequency or time grids; instead, their values are computed on-the-fly at arbitrary complex arguments, allowing precise and flexible evaluation as required by different solvers or output formats. Second, memory usage is optimized by avoiding redundant storage. The internal design separates spectral data and evaluation logic, allowing minimal overhead during the object construction and easing post-processing analysis. 
We recognize that the use of the term "high-performance" could be misleading and rephrases the paragraph on page 16 in a more precise way as

"*The module ${\tt ED\_GFMATRIX}$ implements the class ${\tt gfmatrix}$, which provides an efficient and flexible representation of DCFs. It is structured as a hierarchical, or {\it multilayered}, data container that organizes spectral weights and poles across three levels: (i) the initial eigenstate $|n\rangle$ of the Hamiltonian spectrum, (ii) the amplitude channel associated with this state, corresponding to the applied operator, and (iii) the set of excitations from $|n\rangle$, each characterized by a pole and its corresponding weight.  
This structure allows the ${\tt gfmatrix}$ class to simultaneously handle multiple initial states and operators, enabling on-the-fly evaluation of DCFs at arbitrary complex frequencies $z \in \mathbb{C}$. By avoiding precomputed grids and redundant storage this approach minimizes memory footprint and computational overhead. Through the use of ${\tt gfmatrix}$ EDIpack achieves efficient memory management, consistent storage of spectral data and high-performance post-processing.*"

@@

In the revised manuscript we have corrected this issue. We now use the symbol $Z$ to indicate the exact partition function and reserve the symbol $\tilde{Z}$ to explicitly indicate the partition function approximated using only the finite set of low-energy eigenstates retained in the trace, i.e., those contributing significantly due to their large Boltzmann weights. This makes the approximation consistent on both sides of the involved equations, namely Eqs. (13), (14) and (18), properly reflecting the finite-state summation used in practice.

> *Are the code listings on p. 25 and p. 26 of sufficient interest to readers?*

We acknowledge the referee observation. We believe that the code listings on pages 25 and 26, i.e. the implementation of the fast-algorithm for the evaluation of the impurity RDM,    provide valuable insight into the practical aspects of the implementation discussed in the text, particularly with respect to the subtle issues related to state decomposition and the treatment of fermionic sign conventions where required.

These listings are intended to complement the discussion by offering concrete examples that clarify how the formal logic is realized in code. Given the potential for confusion when dealing with basis state reconstruction and operator ordering in fermionic systems, we believe that these examples can be of interest and practical benefit to users aiming to extend or interface with the library.

@@

> *p. 28: "as the nonsu2 and superc diagonalizations entail nontrivial subtleties in optimizing the off-diagonal components of X". What are these subtleties?*

We acknowledge that our brief comment reads a bit obscure while this point requires a more detailed, yet brief, discussion.

The bath optimization is a crucial step in the DMFT-ED algorithm. Hence, to ensure EDIpack has a large degree of versatility while preserving robustness, we introduced several control parameters for the conjugate-gradient fit. A key one is the definition of the norm, or distance, between matrix functions appearing in Eq. 30.

@@

We thank the referee for carefully reviewing this section and pointing this out. Indeed, the variable ${\tt Wband}$ should not appear in the code listing, as it corresponds to ${\tt D}$, which is correctly defined earlier in the script.  Additionally, the variable ${\tt de}$ represents the energy grid spacing, computed as ${\tt de = 2D / Le}$, where Le is the number of energy points discretizing the Bethe lattice density of states. We have corrected the code listing accordingly to eliminate Wband and clarify the definition of de.

> *p. 42: "provides access to well-tested functions". This is unclear. Functions doing what?*

We thank the referee for pointing out the ambiguity in this sentence. Our intention was to clarify that, although the bath optimization problem is conceptually distinct from the core impurity diagonalization task, for which EDIpack was originally developed, nonetheless the package provides access to a well-tested implementation of the conjugate-gradient fitting procedure for the bath optimization. We have revised the sentence accordingly in the manuscript to make this connection and the role of these functions more explicit by writing:

*"Given the critical importance of this step, EDIpack provides access to a well-tested implementation of the conjugate gradient method for performing the bath optimization, ensuring stability and reproducibility of the results. This task is conceptually distinct from the  diagonalization of the impurity problem, which remains the core focus of the package."*

@@

> *p. 47: generate_kgrid is confusingly complex. There must be a simpler way to accomplish this.*

We agree with the referee that the implementation of ${\tt generate_kgrid}$ could be simplified for clarity. The purpose of this function is merely to construct a uniform 2D grid in the Brillouin zone using the reciprocal lattice vectors. In the case of the square lattice discussed in this example, this implementation is indeed rather overcomplicated, given the orthodonality of the reciprocal lattice vectors. Being completely generic, however, this function can also be used for less simple 2d lattices (e.g. triangular). We further note that in the way it is written, relying entirely on numpy functions, it avoids making use of explicit python ${\tt for}$ loops, which cause performance degradation when the requested number of k-points is large.

More generally, we note that this step is not essential to the core functionalities provided by EDIpack, as it pertains only to the sampling of the Brillouin zone. For more involved tight-binding constructions, users may wish to rely on specialized libraries such as PythTB, which are designed for this purpose.

### 2025-07-17 03_edipack2.tex
Minimal modifications (LC)

From a computational perspective, constructing a symmetry
sector amounts to defining an injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF_e$ that relates the consecutive indexing of the electronic states  $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the Fock states $\ket{I_e}$ in the electronic Fock space in their number representation~\cite{Wallerberger2022PRR}. 
This map is typically implemented as a rank-1 integer array, whose size corresponds to the dimension of the sector $D_{\SS_{\vec{Q}}}$. 
The global Fock state $\ket{I}=\ket{I_e}\ket{p}$, including the phonon part, is univocally determined by the relation: $I\to I_e + p\,2^{2 N_s}$.

@@

Alternatively, setting {\tt ED\_USE\_KANAMORI=F} and {\tt ED\_READ\_UMATRIX=T}, a list of two-body interaction operators can be supplied by the user in a properly
formatted text file, whose name is specified by the input variable {\tt UMATRIX\_FILE}.  
The file must have the suffix {\tt <UMATRIX\_FILE>.re\-star\-t} and should be formatted as follows: 
\begin{lstlisting}[style=mybash,numbers=none]
<NORB> BANDS
i1 j1 k1 l1 U_i1j1k1l1
i2 j2 k2 l2 U_i2j2k2l2
...
\end{lstlisting}
where {\tt NORB} is the number of orbitals $N_\a$.
Each line defines a two-body second-quantized operator of the form 
$$
\hat{H}^{\mathrm{int}}=\frac{1}{2}d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.
$$
where the indices $\{ijkl\}$ refer to the combined spin and internal (orbital, valley...) degrees of freedom, e.g. $i=\alpha\sigma$. 
Note that the coefficient indices $l$ and $k$ are swapped in the input file with respect to the two-body operator they implement. The $\tfrac{1}{2}$ prefactor is applied internally by the parsing routine and need not be accounted for by the user.
Empty lines and lines starting with {\tt \#, \%, !} are ignored.

@@

The module {\tt ED\_GFMATRIX} implements the class {\tt gfmatrix}, which provides an efficient and flexible representation of DCFs. It is structured as a hierarchical, or {\it multi-layered}, data container that organizes spectral weights and poles across three levels: (i) the initial eigenstate $\ket{n}$ of the Hamiltonian spectrum, (ii) the amplitude channel associated with this state, corresponding to the applied operator, and (iii) the set of excitations from $\ket{n}$, each characterized by a pole and its corresponding weight.
This structure allows the {\tt gfmatrix} class to simultaneously handle multiple initial states and operators, enabling on-the-fly evaluation of DCFs at arbitrary complex frequencies $z \in \mathbb{C}$. By avoiding precomputed grids and redundant storage this approach minimizes memory footprint and computational overhead. Through the use of ${\tt gfmatrix}$ EDIpack achieve and efficient memory management, consistent storage of spectral data, and high-performance post-processing.

>>>>

From a computational perspective, constructing a symmetry
sector amounts to defining an injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF_e$ that relates the consecutive indexing of the electronic states  $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the Fock states $\ket{I_e}$ in the electronic Fock space in occupation number representation~\cite{Wallerberger2022PRR}. 
This map is typically implemented as a rank-1 integer array, whose size corresponds to the dimension of the sector $D_{\SS_{\vec{Q}}}$. 
The global Fock state $\ket{I}=\ket{I_e}\ket{p}$, including the phonon part, is univocally determined by the relation: $I\to I_e + p\,2^{2 N_s}$.

@@

Alternatively, setting {\tt ED\_USE\_KANAMORI=F} and {\tt ED\_READ\_UMATRIX=T}, a list of two-body interaction operators can be supplied by the user in a properly
formatted text file, whose name is specified by the input variable {\tt UMATRIX\_FILE}.  
The file must have the suffix {\tt <UMATRIX\_FILE>.re\-star\-t} and should be formatted as follows: 
\begin{lstlisting}[style=mybash,numbers=none]
<NORB> BANDS
i1 j1 k1 l1 U_i1j1k1l1
i2 j2 k2 l2 U_i2j2k2l2
...
\end{lstlisting}
where {\tt NORB} is the number of orbitals $N_\a$ and $\{ijkl\}$ refer to the combined spin ({\tt u/d}) and internal (e.g. orbital index) degrees of freedom. For example, if {\tt i1} refers to orbital 1, spin up, it is written explicitly as {\tt 1 u}.
Each line defines a two-body second-quantized operator of the form 
$$
\hat{H}^{\mathrm{int}}=\frac{1}{2}d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.
$$

Note that the coefficient indices $l$ and $k$ are swapped in the input file with respect to the two-body operator they implement. The $\tfrac{1}{2}$ prefactor is applied internally by the parsing routine and need not be accounted for by the user.
Empty lines and lines starting with {\tt \#, \%, !} are ignored.

@@

The module {\tt ED\_GFMATRIX} implements the class {\tt gfmatrix}, which provides an efficient and flexible representation of DCFs. It is structured as a hierarchical, or {\it multi-layered}, data container that organizes spectral weights and poles across three levels: (i) the initial eigenstate $\ket{n}$ of the Hamiltonian spectrum, (ii) the amplitude channel associated with this state, corresponding to the applied operator, and (iii) the set of excitations from $\ket{n}$, each characterized by a pole and its corresponding weight.
This structure allows the {\tt gfmatrix} class to simultaneously handle multiple initial states and operators, enabling on-the-fly evaluation of DCFs at arbitrary complex frequencies $z \in \mathbb{C}$. By avoiding precomputed grids and redundant storage this approach minimizes memory footprint and computational overhead. Through the use of ${\tt gfmatrix}$ EDIpack achieves efficient memory management, consistent storage of spectral data and high-performance post-processing.

### 2025-07-17 05_examples.tex
Minimal modifications (LC)

The first two steps are handled directly by \NAME routines, while during the subsequent steps the 
 control returns to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the critical importance of this step, \NAME provides access to a well-tested implementation of the conjugate gradient method for performing the bath optimization, ensuring stability and reproducibility of the results, even though this task remains conceptually distinct from the  diagonalization of the impurity problem, which remains the core focus of the package.
Alternative optimization methods can also be employed as 
needed \cite{Mejuto-Zaera2020PRB,Huang2023PRB,Nakatsukasa2018SJOSC}. 
An example of implementation is provided in the following listings.

>>>>

The first two steps are handled directly by \NAME routines, while during the subsequent steps the 
 control returns to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the critical importance of this step, \NAME provides access to a well-tested implementation of the conjugate gradient method for performing the bath optimization, ensuring stability and reproducibility of the results. This task is conceptually distinct from the  diagonalization of the impurity problem, which remains the core focus of the package.
Alternative optimization methods can also be employed as 
needed \cite{Mejuto-Zaera2020PRB,Huang2023PRB,Nakatsukasa2018SJOSC}. 
An example of implementation is provided in the following listings.
